Global Optimization of the Analogue Method by Means of Genetic Algorithms

### 2015-10-15 Optimization - Journal of Climate/template.tex
Changing humidity to moisture.

Multiple variations of the methods exist, and some aspects and parameters will not be detailed hereafter. There are mainly 2 parametrizations that are used most often: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on humidity variables \citep{Obled2002, Bontron2005, Marty2012}.

@@

The other most know parametrization adds a second level of analogy on humidity variables (Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a humidity index made of the product of the precipitable water with the relative humidity at 850~hPa. \cite{Horton2012a} confirmed that this index is better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis considered independently. When adding a second level of analogy, we subsample $N_{2}$ dates in the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When we add a second level of analogy, we keep a higher number of analogues on the first level.

@@

We observed during the calibration of the analogue method that the resulting parameters vary with the initial choices (such as the number of analogues). In addition, the different levels of analogy (eg on the atmospheric circulation and the humidity index) are always calibrated sequentially. However, we can not exclude any dependency between them, which could lead us to select another parametrization if we calibrate them together. Simultaneous calibration of all parameters has never been undertaken so far. Only a global optimization technique can be able to optimize all parameters of all analogy levels simultaneously.

@@

We first optimized the analogy of atmospheric circulation alone, without humidity variables, for our 10 subregions. We started from the most simple analogue method, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. We do not provide the detailed results of the intermediate stages \citep[see][for the details]{Horton2012a}, but only the main conclusions.

@@

\section{Optimization of the analogy with humidity variables}

We previously worked with a method based on a unique level of analogy. However, we know that humidity variables in a second level of analogy do provide improvements to the method (section \ref{section_intro}.\ref{section_analog_method}). So we also wanted to optimize the humidity index, which is a combination of the relative humidity and the precipitable water. In order to do so, we had to introduce a constraint to the optimizer, so that it selects the same temporal window for both variables.

Two methods were assessed: one with 2 humidity predictors (humidity index on 2 pressure levels or at 2 different hours) named z4-hi2, and one with 4 humidity predictors named z4-hi4. When we introduce 2 predictors for the humidity analogy, the number of degrees of freedom raises to 42, and to 54 when we add 4 predictors.

We proceeded to an optimization of both levels of analogy simultaneously. This implies that the analogy of the atmospheric circulation may change due to the new humidity information.

@@

As previously, the optimized parameters differ from one subregion to another, and this event to a greater extent. Detailed examples are still provided for the Swiss Chablais subregion in Table \ref{table:params_GA_z4_hi2} for z4-hi2 and in Table \ref{table:params_GA_z4_hi4} for z4-hi4. The resulting spatial windows are displayed in Figure \ref{fig:spatial_windows_z4-hi2} for z4-hi2, along with the selected pressure levels for the circulation analogy (Table \ref{table:levels_GA_z4_hi2}) and the humidity analogy (Table \ref{table:levels_GA_z4_hi2_H}). Correspondingly, for the method z4-hi4, the spatial windows are shown in Figure \ref{fig:spatial_windows_z4-hi4} and the selected pressure levels are presented in Tables \ref{table:levels_GA_z4_hi4} and \ref{table:levels_GA_z4_hi4_H}.

The CRPSS scores of the optimized methods are provided in Table \ref{table:scores} and amounts to slightly more than 40 \% for both methods and for both periods. This results in a relative improvements that ranges from 11.5 \% to 15.6 \% compared to the reference method on the humidity analogy (Table \ref{table:scores_diff}). As for z4, the z4-hi2 and z4-hi4 methods present improvements of the forecast of significant rainfall (thresholds P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10). The improvement are relatively similar for z4-hi2 (Table \ref{table:scores_thresholds_z4-hi2}) and z4-hi4 (Table \ref{table:scores_thresholds_z4-hi4}), with slightly superior scores for the latter on small precipitation (P\(\geq\)1 mm) and extremes (P\(\geq\)0.5\(\cdot\)P10).

The parameters cross-compatibility has also been assessed for the methods with humidity variables, and are shown in Tables \ref{table:crossing_z4-hi2_calib} and \ref{table:crossing_z4-hi2_valid} for the method z4-hi2, and Tables \ref{table:crossing_z4-hi4_calib} and \ref{table:crossing_z4-hi4_valid} for z4-hi4.

@@

When optimizing a method made of 2 levels of analogy, the introduction of humidity variables in the second level has an influence on the parameter values of the first level. This is first visible on the selection of the pressure levels for the circulation analogy, where the 1000~hPa that was previously systematically selected twice (Table \ref{table:levels_GA_z4}) is now less chosen (once or even not at all) for both za-hi2 (Table \ref{table:levels_GA_z4_hi2}) and za-hi4 (Table \ref{table:levels_GA_z4_hi4}). There is indeed a shift of the previously selected 1000~hPa for higher levels, which is even slightly stronger with 4 humidity predictors than with 2. This change is likely due to the fact that when considering only the circulation analogy, the method tries to take into account information that can serve as proxy for humidity assessment, whereas it doesn't need it with the humidity index. This aspect has never been demonstrated, as classic calibration tools do not allow it. It can only be assessed by a global optimization technique that can tune jointly both levels of analogy.

The selected pressure levels for the analogy on the humidity index are strongly centred around 700~hPa and 600~hPa. No other value has been selected when considering 2 pressure levels (Table \ref{table:levels_GA_z4_hi2_H}), and when considering 4 levels, the 850~hPa and 500~hPa levels are additionally selected. However, even in this latter method, the 700~hPa and 600~hPa levels still hold 78 \% of the selection. It is thus more efficient, in terms of forecasting performance, to consider one of this level multiple times at different hours rather than using another pressure level. Besides, the optimizer never chose the same pressure level at the same hour for any variable, even though it was allowed to do so.

The selection of the temporal windows for the atmospheric circulation is similar to the preceding optimization (in the order of increasing pressure: 12~h, 24/30~h, 12~h, 30~h), but sometimes with a bit more variability. When it comes to the humidity analogy, there is a clear trend to select 12~h and 24~h when considering 2 predictors. The z4-hi4 method presents more variability on the selection of the temporal windows for the humidity index.

The optimized spatial windows for the atmospheric circulation have also changed (Figures \ref{fig:spatial_windows_z4-hi2} and \ref{fig:spatial_windows_z4-hi4}). The very large domains on the upper level of the 4 first subregions are not present anymore, and more variability can be observed. The selected points for the analogy on the humidity index are always located nearby the catchment, including at least 1 of the nearest points from the reanalysis dataset, and the spatial windows are relatively small. Thus, for our case study, there is no need to look for distant humidity information and the search could be reduced to a smaller domain.

The CRPSS scores were improved by considering the humidity information (Table \ref{table:scores}). The optimized methods also perform better than the reference method based on the humidity index (Table \ref{table:scores_diff}). However, there is no drastic difference between both z4-hi2 and z4-hi4 methods, which suggests that considering 4 humidity predictors is not absolutely necessary. When it comes to improvements for days with precipitation above our 3 thresholds (P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10), the conclusion is the same as before, that is a significant improvement of the forecast, mainly for heavy rainfall.

The analysis of the parameters cross-compatibility (Tables \ref{table:crossing_z4-hi2_calib} to \ref{table:crossing_z4-hi4_valid}) is also very similar to the one on the circulation analogy only. The same pattern can be observed, with a drop of performance for the subregions under different meteorological influences. However, the losses of performance are globally more important than before, suggesting that more complex methods with humidity variables are less transposable to another subregion, even though it is located within the same grid of the reanalysis dataset.

@@

\begin{itemize}
	\item Using 4 pressure levels for the analogy of circulation. It seems to be the optimal number for the studied region. Beyond that value the validation score drops, revealing a lost in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogues number, the selection of the pressure levels and the time windows, and the spatial windows. These parameters are highly interdependent, so we need to optimize them jointly in order to identify optimal combinations. Traditional calibration procedures based on a systematic assessment of every combination is no more possible when considering more than 2 pressure levels.
	\item The introduction of distinct spatial windows between pressure levels. Indeed, the synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting rainfall is not necessarily located in the same area from one level to another. The optimized spatial windows are consistent in between the subregions.
	\item The weighting of the analogy criteria between different pressure levels. It can be influenced by the variability of the geopotential with altitude, and the change of some levels significance with the targeted region. There is a trend for the weighting of circulation predictors to decrease with the increase in pressure, as one can see in Figure \ref{fig:levels_weights} for the method z4, and in Figure \ref{fig:levels_weights_average} for the averages over the 3 methods. However, the values stay around equity. This may not be the most influencing factor, and we may suggest to remove it first when trying to reduce the number of degrees of freedom.
	\item The joint optimization of the circulation and humidity analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, we must consider them jointly.
\end{itemize}

@@

We tried to optimize the preselection period (the 4-months window) jointly with the other parameters, but we did not get improvement. We also tried optimizing the humidity flux, which is composed of the humidity index multiplied with the wind flux. However, the results were not better than when we considered the humidity index alone. This may be related to the fact that the optimizer tries to provide the best analogy of the atmospheric circulation in the first place, which makes the wind information less relevant in the second level of analogy.

As we could observe, methods with a higher complexity that integrate humidity predictors are less transposable than simpler ones. We also noticed in another unpublished work, that it is by far better to optimize for 2 subregions jointly rather than to optimize on one and to apply its parametrization to the other. Finally, the creation of the subregions is an important process and should be handled with care. Indeed, the geographical proximity is not always the leading factor to define a subregion. For example, the Southeast ridges subregion do not behave like its surrounding and differ in its parametrization, due to different leading meteorological influences.

@@

The parameters resulting from the optimization are very consistent for the analogy of the atmospheric circulation, in terms of selection of the pressure levels, and the temporal and spatial windows. There are clear trends or even identical results for subregions under similar meteorological influences, which confirm that the optimized parameters are coherent, despite an eventual first impression of a great variability in the spatial windows. When adding humidity variables, the results show a higher variability, but remains highly acceptable and coherent. Indeed, the selection of the pressure level for the humidity index is for example very constant.

@@

We were able to identify that there is a parametric dependence between the analogy of circulation and the moisture humidity. When we consider the two levels together, the optimal parameters of the analogy of circulation are different. This complexity can only be exploited in a suitable manner by global optimization methods.

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Parameters of the reference method with humidity variables.}
	\begin{center}
		\begin{tabular}{ccccc}
			\hline \textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z 1000~hPa & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{70} \\
			& Z 500~hPa & 24~h & & \\ 
			\hline
			\multirow{2}{*}{2} & TPW * RH 850~hPa & 12~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{30} \\
			& TPW * RH 850~hPa & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R2}
\end{table}

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Improvement (\%) of the CRPSS of the three methods. For z4, the improvement is relative to the reference method on the atmospheric circulation. For z4-hi2 and z4-hi4, the improvement is relative to the reference method on the circulation and humidity variables.}
	\begin{center}
		\begin{tabular}{rrarara}
			\hline 
			\textbf{ID} & \multicolumn{2}{c}{\textbf{z4}} & \multicolumn{2}{c}{\textbf{z4-hi2}} & \multicolumn{2}{c}{\textbf{z4-hi4}} \\ 
			\hline 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			\textbf{1} & 13.6 & 13.0 & 11.9 & 9.9 & 12.8 & 9.6 \\ \hline
			\textbf{2} & 13.3 & 7.5 & 11.2 & 9.4 & 12.5 & 10.8 \\ \hline
			\textbf{3} & 12.2 & 9.0 & 13.2 & 12.0 & 14.1 & 13.5 \\ \hline
			\textbf{4} & 14.2 & 10.5 & 12.4 & 10.6 & 14.5 & 12.5 \\ \hline
			\textbf{5} & 14.5 & 8.8 & 12.5 & 9.1 & 14.1 & 10.0 \\ \hline
			\textbf{6} & 13.3 & 8.6 & 11.5 & 8.2 & 14.1 & 8.4 \\ \hline
			\textbf{7} & 22.8 & 10.6 & 19.7 & 15.7 & 21.7 & 16.3 \\ \hline
			\textbf{8} & 20.7 & 11.9 & 19.2 & 13.5 & 20.8 & 14.0 \\ \hline
			\textbf{9} & 14.4 & 9.5 & 15.8 & 15.2 & 16.7 & 14.1 \\ \hline
			\textbf{10} & 14.2 & 9.4 & 12.9 & 11.2 & 14.6 & 11.8 \\ \hline
			\textbf{av.} & \textbf{15.3} & \textbf{9.9} &\textbf{14.0} & \textbf{11.5} & \textbf{15.6} & \textbf{12.1} \\ \hline 
		\end{tabular} 
	\end{center}
	\label{table:scores_diff}
\end{table}

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Parameters of the z4-hi2 method (atmospheric circulation on 4 levels and humidity index on 2 levels), optimized for the Chablais subregion. Same conventions as Table \ref{table:params_GA_z4}.}
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline \textbf{L} & \textbf{V} & \textbf{H} & \textbf{D} & \textbf{W} & \textbf{C} & \textbf{N} \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z300} & \multirow{2}{*}{12~h} & 0.0 - 15.0 \degree E & \multirow{2}{*}{23\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{74} \\
			& & & 42.5 - 55.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z500} & \multirow{2}{*}{30~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{25\%} & & \\ 
			& & & 40.0 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z850} & \multirow{2}{*}{12~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 0.0 - 20.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 40.0 - 47.5 \degree N & & & \\ 
			\hline 
			\multirow{4}{*}{2} & TPW & \multirow{2}{*}{12~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{57\%} & \multirow{4}{*}{RMSE} & \multirow{4}{*}{25} \\
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			& TPW & \multirow{2}{*}{24~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{43\%} & & \\ 
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_z4_hi2}
\end{table}

>>>>

Multiple variations of the methods exist, and some aspects and parameters will not be detailed hereafter. There are mainly 2 parametrizations that are used most often: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

@@

The other most know parametrization adds a second level of analogy on moisture variables (Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the precipitable water with the relative humidity at 850~hPa. \cite{Horton2012a} confirmed that this index is better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis considered independently. When adding a second level of analogy, we subsample $N_{2}$ dates in the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When we add a second level of analogy, we keep a higher number of analogues on the first level.

@@

We observed during the calibration of the analogue method that the resulting parameters vary with the initial choices (such as the number of analogues). In addition, the different levels of analogy (eg on the atmospheric circulation and the moisture index) are always calibrated sequentially. However, we can not exclude any dependency between them, which could lead us to select another parametrization if we calibrate them together. Simultaneous calibration of all parameters has never been undertaken so far. Only a global optimization technique can be able to optimize all parameters of all analogy levels simultaneously.

@@

We first optimized the analogy of atmospheric circulation alone, without moisture variables, for our 10 subregions. We started from the most simple analogue method, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. We do not provide the detailed results of the intermediate stages \citep[see][for the details]{Horton2012a}, but only the main conclusions.

@@

\section{Optimization of the analogy with moisture variables}

We previously worked with a method based on a unique level of analogy. However, we know that moisture variables in a second level of analogy do provide improvements to the method (section \ref{section_intro}.\ref{section_analog_method}). So we also wanted to optimize the moisture index, which is a combination of the relative humidity and the precipitable water. In order to do so, we had to introduce a constraint to the optimizer, so that it selects the same temporal window for both variables.

Two methods were assessed: one with 2 moisture predictors (moisture index on 2 pressure levels or at 2 different hours) named z4-hi2, and one with 4 moisture predictors named z4-hi4. When we introduce 2 predictors for the moisture analogy, the number of degrees of freedom raises to 42, and to 54 when we add 4 predictors.

We proceeded to an optimization of both levels of analogy simultaneously. This implies that the analogy of the atmospheric circulation may change due to the new moisture information.

@@

As previously, the optimized parameters differ from one subregion to another, and this event to a greater extent. Detailed examples are still provided for the Swiss Chablais subregion in Table \ref{table:params_GA_z4_hi2} for z4-hi2 and in Table \ref{table:params_GA_z4_hi4} for z4-hi4. The resulting spatial windows are displayed in Figure \ref{fig:spatial_windows_z4-hi2} for z4-hi2, along with the selected pressure levels for the circulation analogy (Table \ref{table:levels_GA_z4_hi2}) and the moisture analogy (Table \ref{table:levels_GA_z4_hi2_H}). Correspondingly, for the method z4-hi4, the spatial windows are shown in Figure \ref{fig:spatial_windows_z4-hi4} and the selected pressure levels are presented in Tables \ref{table:levels_GA_z4_hi4} and \ref{table:levels_GA_z4_hi4_H}.

The CRPSS scores of the optimized methods are provided in Table \ref{table:scores} and amounts to slightly more than 40 \% for both methods and for both periods. This results in a relative improvements that ranges from 11.5 \% to 15.6 \% compared to the reference method on the moisture analogy (Table \ref{table:scores_diff}). As for z4, the z4-hi2 and z4-hi4 methods present improvements of the forecast of significant rainfall (thresholds P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10). The improvement are relatively similar for z4-hi2 (Table \ref{table:scores_thresholds_z4-hi2}) and z4-hi4 (Table \ref{table:scores_thresholds_z4-hi4}), with slightly superior scores for the latter on small precipitation (P\(\geq\)1 mm) and extremes (P\(\geq\)0.5\(\cdot\)P10).

The parameters cross-compatibility has also been assessed for the methods with moisture variables, and are shown in Tables \ref{table:crossing_z4-hi2_calib} and \ref{table:crossing_z4-hi2_valid} for the method z4-hi2, and Tables \ref{table:crossing_z4-hi4_calib} and \ref{table:crossing_z4-hi4_valid} for z4-hi4.

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level has an influence on the parameter values of the first level. This is first visible on the selection of the pressure levels for the circulation analogy, where the 1000~hPa that was previously systematically selected twice (Table \ref{table:levels_GA_z4}) is now less chosen (once or even not at all) for both za-hi2 (Table \ref{table:levels_GA_z4_hi2}) and za-hi4 (Table \ref{table:levels_GA_z4_hi4}). There is indeed a shift of the previously selected 1000~hPa for higher levels, which is even slightly stronger with 4 moisture predictors than with 2. This change is likely due to the fact that when considering only the circulation analogy, the method tries to take into account information that can serve as proxy for moisture assessment, whereas it doesn't need it with the moisture index. This aspect has never been demonstrated, as classic calibration tools do not allow it. It can only be assessed by a global optimization technique that can tune jointly both levels of analogy.

The selected pressure levels for the analogy on the moisture index are strongly centred around 700~hPa and 600~hPa. No other value has been selected when considering 2 pressure levels (Table \ref{table:levels_GA_z4_hi2_H}), and when considering 4 levels, the 850~hPa and 500~hPa levels are additionally selected. However, even in this latter method, the 700~hPa and 600~hPa levels still hold 78 \% of the selection. It is thus more efficient, in terms of forecasting performance, to consider one of this level multiple times at different hours rather than using another pressure level. Besides, the optimizer never chose the same pressure level at the same hour for any variable, even though it was allowed to do so.

The selection of the temporal windows for the atmospheric circulation is similar to the preceding optimization (in the order of increasing pressure: 12~h, 24/30~h, 12~h, 30~h), but sometimes with a bit more variability. When it comes to the moisture analogy, there is a clear trend to select 12~h and 24~h when considering 2 predictors. The z4-hi4 method presents more variability on the selection of the temporal windows for the moisture index.

The optimized spatial windows for the atmospheric circulation have also changed (Figures \ref{fig:spatial_windows_z4-hi2} and \ref{fig:spatial_windows_z4-hi4}). The very large domains on the upper level of the 4 first subregions are not present anymore, and more variability can be observed. The selected points for the analogy on the moisture index are always located nearby the catchment, including at least 1 of the nearest points from the reanalysis dataset, and the spatial windows are relatively small. Thus, for our case study, there is no need to look for distant moisture information and the search could be reduced to a smaller domain.

The CRPSS scores were improved by considering the moisture information (Table \ref{table:scores}). The optimized methods also perform better than the reference method based on the moisture index (Table \ref{table:scores_diff}). However, there is no drastic difference between both z4-hi2 and z4-hi4 methods, which suggests that considering 4 moisture predictors is not absolutely necessary. When it comes to improvements for days with precipitation above our 3 thresholds (P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10), the conclusion is the same as before, that is a significant improvement of the forecast, mainly for heavy rainfall.

The analysis of the parameters cross-compatibility (Tables \ref{table:crossing_z4-hi2_calib} to \ref{table:crossing_z4-hi4_valid}) is also very similar to the one on the circulation analogy only. The same pattern can be observed, with a drop of performance for the subregions under different meteorological influences. However, the losses of performance are globally more important than before, suggesting that more complex methods with moisture variables are less transposable to another subregion, even though it is located within the same grid of the reanalysis dataset.

@@

\begin{itemize}
	\item Using 4 pressure levels for the analogy of circulation. It seems to be the optimal number for the studied region. Beyond that value the validation score drops, revealing a lost in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogues number, the selection of the pressure levels and the time windows, and the spatial windows. These parameters are highly interdependent, so we need to optimize them jointly in order to identify optimal combinations. Traditional calibration procedures based on a systematic assessment of every combination is no more possible when considering more than 2 pressure levels.
	\item The introduction of distinct spatial windows between pressure levels. Indeed, the synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting rainfall is not necessarily located in the same area from one level to another. The optimized spatial windows are consistent in between the subregions.
	\item The weighting of the analogy criteria between different pressure levels. It can be influenced by the variability of the geopotential with altitude, and the change of some levels significance with the targeted region. There is a trend for the weighting of circulation predictors to decrease with the increase in pressure, as one can see in Figure \ref{fig:levels_weights} for the method z4, and in Figure \ref{fig:levels_weights_average} for the averages over the 3 methods. However, the values stay around equity. This may not be the most influencing factor, and we may suggest to remove it first when trying to reduce the number of degrees of freedom.
	\item The joint optimization of the circulation and moisture analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, we must consider them jointly.
\end{itemize}

@@

We tried to optimize the preselection period (the 4-months window) jointly with the other parameters, but we did not get improvement. We also tried optimizing the moisture flux, which is composed of the moisture index multiplied with the wind flux. However, the results were not better than when we considered the moisture index alone. This may be related to the fact that the optimizer tries to provide the best analogy of the atmospheric circulation in the first place, which makes the wind information less relevant in the second level of analogy.

As we could observe, methods with a higher complexity that integrate moisture predictors are less transposable than simpler ones. We also noticed in another unpublished work, that it is by far better to optimize for 2 subregions jointly rather than to optimize on one and to apply its parametrization to the other. Finally, the creation of the subregions is an important process and should be handled with care. Indeed, the geographical proximity is not always the leading factor to define a subregion. For example, the Southeast ridges subregion do not behave like its surrounding and differ in its parametrization, due to different leading meteorological influences.

@@

The parameters resulting from the optimization are very consistent for the analogy of the atmospheric circulation, in terms of selection of the pressure levels, and the temporal and spatial windows. There are clear trends or even identical results for subregions under similar meteorological influences, which confirm that the optimized parameters are coherent, despite an eventual first impression of a great variability in the spatial windows. When adding moisture variables, the results show a higher variability, but remains highly acceptable and coherent. Indeed, the selection of the pressure level for the moisture index is for example very constant.

@@

We were able to identify that there is a parametric dependence between the analogy of circulation and the moisture. When we consider the two levels together, the optimal parameters of the analogy of circulation are different. This complexity can only be exploited in a suitable manner by global optimization methods.

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Parameters of the reference method with moisture variables.}
	\begin{center}
		\begin{tabular}{ccccc}
			\hline \textbf{Level} & \textbf{Variable} & \textbf{Hour} & \textbf{Criteria} & \textbf{Nb} \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z 1000~hPa & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{70} \\
			& Z 500~hPa & 24~h & & \\ 
			\hline
			\multirow{2}{*}{2} & TPW * RH 850~hPa & 12~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{30} \\
			& TPW * RH 850~hPa & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R2}
\end{table}

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Improvement (\%) of the CRPSS of the three methods. For z4, the improvement is relative to the reference method on the atmospheric circulation. For z4-hi2 and z4-hi4, the improvement is relative to the reference method on the circulation and moisture variables.}
	\begin{center}
		\begin{tabular}{rrarara}
			\hline 
			\textbf{ID} & \multicolumn{2}{c}{\textbf{z4}} & \multicolumn{2}{c}{\textbf{z4-hi2}} & \multicolumn{2}{c}{\textbf{z4-hi4}} \\ 
			\hline 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			\textbf{1} & 13.6 & 13.0 & 11.9 & 9.9 & 12.8 & 9.6 \\ \hline
			\textbf{2} & 13.3 & 7.5 & 11.2 & 9.4 & 12.5 & 10.8 \\ \hline
			\textbf{3} & 12.2 & 9.0 & 13.2 & 12.0 & 14.1 & 13.5 \\ \hline
			\textbf{4} & 14.2 & 10.5 & 12.4 & 10.6 & 14.5 & 12.5 \\ \hline
			\textbf{5} & 14.5 & 8.8 & 12.5 & 9.1 & 14.1 & 10.0 \\ \hline
			\textbf{6} & 13.3 & 8.6 & 11.5 & 8.2 & 14.1 & 8.4 \\ \hline
			\textbf{7} & 22.8 & 10.6 & 19.7 & 15.7 & 21.7 & 16.3 \\ \hline
			\textbf{8} & 20.7 & 11.9 & 19.2 & 13.5 & 20.8 & 14.0 \\ \hline
			\textbf{9} & 14.4 & 9.5 & 15.8 & 15.2 & 16.7 & 14.1 \\ \hline
			\textbf{10} & 14.2 & 9.4 & 12.9 & 11.2 & 14.6 & 11.8 \\ \hline
			\textbf{av.} & \textbf{15.3} & \textbf{9.9} &\textbf{14.0} & \textbf{11.5} & \textbf{15.6} & \textbf{12.1} \\ \hline 
		\end{tabular} 
	\end{center}
	\label{table:scores_diff}
\end{table}

@@

\begin{table}[htbp]
	\footnotesize
	\caption{Parameters of the z4-hi2 method (atmospheric circulation on 4 levels and moisture index on 2 levels), optimized for the Chablais subregion. Same conventions as Table \ref{table:params_GA_z4}.}
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline \textbf{L} & \textbf{V} & \textbf{H} & \textbf{D} & \textbf{W} & \textbf{C} & \textbf{N} \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z300} & \multirow{2}{*}{12~h} & 0.0 - 15.0 \degree E & \multirow{2}{*}{23\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{74} \\
			& & & 42.5 - 55.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z500} & \multirow{2}{*}{30~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{25\%} & & \\ 
			& & & 40.0 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z850} & \multirow{2}{*}{12~h} & 0.0 - 10.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 0.0 - 20.0 \degree E & \multirow{2}{*}{26\%} & & \\ 
			& & & 40.0 - 47.5 \degree N & & & \\ 
			\hline 
			\multirow{4}{*}{2} & TPW & \multirow{2}{*}{12~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{57\%} & \multirow{4}{*}{RMSE} & \multirow{4}{*}{25} \\
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			& TPW & \multirow{2}{*}{24~h} & 5.0 - 10.0 \degree E & \multirow{2}{*}{43\%} & & \\ 
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_z4_hi2}
\end{table}

\begin{table}[htbp]
	\footnotesize
	\caption{Parameters of the z4-hi4 method (atmospheric circulation on 4 levels and moisture index on 4 levels), optimized for the Chablais subregion. Same conventions as Table \ref{table:params_GA_z4}.}
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline \textbf{L} & \textbf{V} & \textbf{H} & \textbf{D} & \textbf{W} & \textbf{C} & \textbf{N} \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z400} & \multirow{2}{*}{6~h} & -5.0 - 12.5 \degree E & \multirow{2}{*}{20\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{74} \\
			& & & 45.0 - 52.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z500} & \multirow{2}{*}{30~h} & 0.0 - 12.5 \degree E & \multirow{2}{*}{28\%} & & \\ 
			& & & 40.0 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z850} & \multirow{2}{*}{12~h} & 0.0 - 12.5 \degree E & \multirow{2}{*}{28\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 0.0 - 20.0 \degree E & \multirow{2}{*}{24\%} & & \\ 
			& & & 40.0 - 47.5 \degree N & & & \\ 
			\hline 
			\multirow{8}{*}{2} & TPW & \multirow{2}{*}{24~h} & 7.5 - 10.0 \degree E & \multirow{2}{*}{27\%} & \multirow{8}{*}{RMSE} & \multirow{8}{*}{25} \\
			& *RH500 & & 45.0 - 47.5 \degree N & & & \\ 
			& TPW & \multirow{2}{*}{12~h} & 5.0 - 7.5 \degree E & \multirow{2}{*}{41\%} & & \\ 
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			& TPW & \multirow{2}{*}{18~h} & 0.0 - 7.5 \degree E & \multirow{2}{*}{19\%} & & \\ 
			& *RH700 & & 45.0 - 47.5 \degree N & & & \\ 
			& TPW & \multirow{2}{*}{24~h} & 2.5 - 15.0 \degree E & \multirow{2}{*}{13\%} & & \\ 
			& *RH850 & & 45.0 - 47.5 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_z4_hi4}
\end{table}
